---
title: "Hands-on Exercise 1"
author: "Bao Jie Yeo"
date: '2024-08-16'
execute: 
  eval: false
  echo: true
  freeze: true
---

## Getting Started

```{r}
pacman::p_load(sf, tidyverse)
```

## Import Geospatial Data

Use `st_read` for loading and parsing.

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
cyclingpath <- st_read(dsn = "data/geospatial", 
                       layer = "CyclingPathGazette")
```

Provide *entire filepath* for KML file.

```{r}
preschool <- st_read("data/geospatial/PreSchoolsLocation.kml")
```

## Checking Content of Simple Data Frame

`st_geometry` prints basic information of the feature class

```{r}
st_geometry(mpsz)
```

`glimpse` provides info on the data type of each fields

```{r}
glimpse(mpsz)
```

`head` shows the complete info of a feature object

```{r}
head(mpsz, n=5)
```

## Plotting Geospatial Data

`plot` all attributes of mpsz

```{r}
plot(mpsz)
```

`plot` only one attribute
```{r}
plot(mpsz["PLN_AREA_N"])
```

`plot` only geometry
```{r}
plot(st_geometry(mpsz))
```

## Coordinate Systems
```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
```

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## Working with Aspatial Data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings)
```

```{r}
listings_sf <- st_as_sf(listings, 
                        coords = c("longitude", "latitude"),
                        crs=4326) |>
  st_transform(crs = 3414)
glimpse(listings_sf)
```

## Geoprocessing with sf

### Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                            dist=5, nQuadSegs = 30)
```

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

### Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
```

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

```{r}
mpsz3414$Area <- mpsz3414 |>
  st_area()
mpsz3414 <- mpsz3414 |>
  mutate(`PreSch Density` = `PreSch Count` / Area * 1000000)
```

## Data Visualizations

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x = as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle = "There are many planning sub-zones with a single pre-school, on the other hand,   \nthere are two planning sub-zones with at least 20 pre-schools",
       x = "Pre-school density (per km sq)",
       y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y =`PreSch Count`, 
           x = as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue",
             na.rm=TRUE) +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
       x = "Pre-school density (per km sq)",
       y = "Pre-school count")
```





