---
title: "In Class Exercise 1"
author: "Bao Jie Yeo"
date: '2024-08-19'
---

## Load relevant packages

```{r}
pacman::p_load(sf, tidyverse)
```

## Exploring the MPSZ-2019 Dataset

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MPSZ-2019")
```

::: callout-note
### Reflection

The imported data contains 332 features (representing different subzones in SG) with 6 fields. The geometry type is MULTIPOLYGON, with some areas consisting of multiple polygons (e.g. islands),
:::

## Check Content of Data Frame

First, we print the information of the feature class using `st_geometry`

```{r}
st_geometry(mpsz)
```

We can use `glimpse` to provide information on the data type of each field

```{r}
glimpse(mpsz)
```

Finally, use `head` to show the complete info of a feature object

```{r}
head(mpsz, n=5)
```

Python users will be familiar with `tail` to show the last few rows of a dataframe

```{r}
tail(mpsz, n=5)
```

## Plotting Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(mpsz["PLN_AREA_C"])
```

```{r}
plot(st_geometry(mpsz))
```

## Coordinate Systems

```{r}
st_crs(mpsz)
```

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
st_crs(mpsz3414)
```

## Geoprocessing with sf library

### Buffering

```{r}
st_crs(mpsz)  # Check current CRS
mpsz_projected <- st_transform(mpsz, 3414)  # Transform to Singapore SVY21 (EPSG:3414)
mpsz_buffer <- st_buffer(mpsz_projected, dist = 5, nQuadSegs = 30)
```
