---
title: "In Class Exercise 10"
author: "Bao Jie Yeo"
date: '2024-10-28'
date-modified: '2024-10-28'
execute: 
  eval: true
  echo: true
  freeze: true
---

## Loading Packages

```{r}
pacman::p_load(olsrr, ggstatsplot, sf, 
               tmap, tidyverse, gtsummary,
               performance, see, sfdep)
```

## Importing Data

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")

mpsz <- read_rds("data/rds/mpsz.rds")

condo_resale_sf <- read_rds(
  "data/rds/condo_resale_sf.rds")
```

## Correlation Analysis: ggstatsplot

```{r}
#| fig-width: 13
#| fig-height: 8
ggcorrmat(condo_resale[, 5:23])
```

## Building Hedonic Price Model using Multiple Linear Regression

```{r}
condo_mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + 
                  AGE	+ PROX_CBD + PROX_CHILDCARE + 
                  PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA + 
                  PROX_HAWKER_MARKET	+ PROX_KINDERGARTEN	+ 
                  PROX_MRT	+ PROX_PARK	+ PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL	+ 
                  PROX_SUPERMARKET + PROX_BUS_STOP + 
                  NO_Of_UNITS + FAMILY_FRIENDLY + 
                  FREEHOLD + LEASEHOLD_99YR, 
                data=condo_resale_sf)
summary(condo_mlr)
```

## Generating tidy linear regression report

```{r}
ols_regress(condo_mlr)
```

## Variable Selection

```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE)
condo_fw_mlr
```

```{r}
#| fig-width: 12
#| fig-height: 10
plot(condo_fw_mlr)
```

## Backward stepwise multiple linear regression

```{r}
condo_bw_mlr <- ols_step_backward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE)
condo_bw_mlr
```

## Stepwise multiple linear regression

```{r}
condo_sb_mlr <- ols_step_both_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE)
condo_sb_mlr
```

## Comparing Performance

```{r}
metric <- compare_performance(condo_mlr, 
                    condo_fw_mlr$model,
                    condo_bw_mlr$model,
                    condo_sb_mlr$model)
```

```{r}
metric$Name <- gsub(".*\\\\([a-zA-Z0-9_]+)\\\\, \\\\model\\\\.*", "\\1", metric$Name)
```

### Visual Comparison

```{r}
plot(metric)
```

### Visualizing Model Parameters

```{r}
ggcoefstats(condo_sb_mlr$model,
            sort = "ascending")
```

## Checking for Multicollinearity

```{r}
check_collinearity(condo_sb_mlr$model)
```

### Visualizing Multicollinearity

```{r}
#| fig-width: 12
#| fig-height: 6
plot(check_collinearity(condo_sb_mlr$model)) +
  theme(axis.text.x = element_text(
    angle = 45, hjust = 1))
```

## Linearity Assumption Test

```{r}
out <- plot(check_model(condo_sb_mlr$model, 
                        panel = FALSE))
out[[2]]
```

## Normality Assumption Test

```{r}
plot(check_normality(condo_sb_mlr$model))
```

## Checking Outliers

```{r}
outliers <- check_outliers(condo_sb_mlr$model,
                           method = "cook")
outliers
```

### Visualizing Outliers

```{r}
plot(check_outliers(condo_sb_mlr$model,
                           method = "cook"))
```

## Spatial Non-stationary Assumption

The hedonic model we try to build are using geographically referenced attributes, hence it is also important for us to visual the residual of the hedonic pricing model.

-   Ho: The residuals are randomly distributed (also known as spatial stationary) 
-   H1: The residuals are spatially non-stationary

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>%
  rename(`FW_MLR_RES` = `condo_fw_mlr$model$residuals`)
```

```{r}
condo_resale_sf <- cbind(condo_resale_sf, 
                        mlr_output$FW_MLR_RES) %>%
  rename(`MLR_RES` = `mlr_output.FW_MLR_RES`)
```

```{r}
tmap_mode("plot")
tm_shape(mpsz)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale_sf) +  
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style="quantile") 
```

## Spatial Stationary Test

```{r}
condo_resale_sf <- condo_resale_sf %>%
  mutate(nb = st_knn(geometry, k=6,
                     longlat = FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale_sf$MLR_RES, 
                  condo_resale_sf$nb, 
                  condo_resale_sf$wt, 
                  alternative = "two.sided", 
                  nsim = 99)
```

### Statistical Interpretation

The Global Moran's I test for residual spatial autocorrelation shows that **its p-value is less than the alpha value of 0.05**. Hence, we **reject the null hypothesis** that the residuals are randomly distributed.

Since the Observed Global Moran I = 0.25586 which is greater than 0, we can infer that the residuals resemble **cluster distribution**.